### 2012-10-25 3.tex
chapter 3 draft is done.

There is a long tradition of distinguishing between structural phonotactic gaps and those which are simply accidental \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}. In purely statistical terms, an accidental gap is one which is likely to arise without any antecedent cause, i.e., by chance. The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps in mentalism: an accidental gap is one which which lacks a mental antecedent because it goes unlearned by speakers.

Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}. The assumption appears to be that statistically valid static co-occurrence restrictions will be reflected by wordlikeness judgements. While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}. In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:

@@

There are several arguments against this inference. First, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality. One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}), who recognizes the a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}, and that Classical Latin intervocalic \emph{s} is rare as a result. However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.} Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} 'fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day. \citet{Iverson2005} label this constraint ``phonologically accidental'', since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and borrowings (e.g., \emph{douche}, \emph{capiche}) in utter disregard of this generalization. Secondly, wordlikeness generalizations are absolutely constrained by negative results in learnability (see \citealt{Yang2012} for a recent review). Finally, there are a number of results which suggest patterns which were obvious to the linguist go unlearned \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}. This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations. The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may overstate the phonotactic generalizations internalized by speakers.

This chapter is a comprehensive investigation of three well-known phonotactic generalizations in Turkish using lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}. \citeauthor{Zimmer1969}'s experiments merit reconsideration because there is widespread disagreement concerning the results and prior discussions do not relate behavioral data to lexical statistics and to competing formalizations of the generalizations involved.

One of these generalizations, \textsc{Labial Attraction}, is shown to be highly statistically reliable but to be inactive in the wordlikeness task. This further undermines the widespread use of statistical criteria to determine which of the vast number of statistically reliable phonotactic generalizations are internalized by speakers and which are not. The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is phonologically inactive, corresponding to no phonological process in the language.

@@

\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences, constraints which are the focus of many subsequent studies. In this section, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence bearing on speakers' knowlege of the restrictions. The following feature specification for the eight vowels of Turkish vowels is assumed throughout.

@@

Two less familiar notations are used in this chapter. First, directional application conditions are used to derive the left-to-right, iterative properties of the harmony rules. Directional application explains provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}). In fact, harmony in Turkish is one such case, insofar as simultaneous application makes the wrong predictions. Further, \citet{Johnson1972} and \citet{Kaplan1994} prove that rules specified for directional application are no more complex in the terms of formal learnability than that do not. Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used \citep{McCawley1973}. A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F].\footnote{This is more restrictive than the Greek-letter variable notation, in that the value of $=$ for some feature F cannot be applied to F$'$ if F $\ne$ F$'$; \citet{Odden2012} dismisses well-known counterexamples against this restriction adduced in \emph{SPE} (352-353).}

@@

If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below. \textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels. For instance, the nominative plural (nom.pl.) suffix is \emph{-ler}, when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].

\begin{example}[The Turkish nominative]
\label{turknom}
\begin{tabular}{lllll}
   & \emph{nom.sg.} & \emph{nom.pl.} \\
a. & {ip}           & {ipler}    & `rope'         & \citep[][216]{Clements1982} \\
%  & {el}           & {eller}    & `hand'         \\
   & {köy}          & {köyler}   & `village'      \\
   & {yüz}          & {yüzler}   & `face'         \\
   & {kız}          & {kızlar}   & `girl'         \\
%  & {sap}          & {saplar}   & `stalk'        \\
%  & {son}          & {sonlar}   & `end'          \\
   & {pul}          & {pullar}   & `stamp'        \\
b. & {neden}        & {nedenler} & `reason'       & \citep{TELL} \\
   & {kiler}        & {kilerler} & `pantry'       \\
   & {pelür}        & {pelürler} & `onionskin'    \\
   & {boğaz}        & {boğazlar} & `throat'       \\
   & {sapık}        & {sapıklar} & `pervert'      \\
\end{tabular}
\end{example}

A few complications arise, however. First, as shown in (\ref{turkexcept}a), not all polysyllabic roots conform to \textsc{Backness Harmony}. In this case, suffix vowels generally exhibit harmony with the final root vowel.

\begin{example}[Exceptional Turkish nominatives] 
\label{turkexcept}
\begin{tabular}{lllll}
   & \emph{nom.sg.} & \emph{nom.pl.}             &                    \\
a. & {mezar}        & {mezarlar} & `grave'       & \citep{TELL}       \\
   & {model}        & {modeller} & `model'                            \\
   & {silah}        & {silahlar} & `weapon'                           \\
   & {memur}        & {memurlar} & `official'                         \\
   & {sabun}        & {sabunlar} & `soap'                             \\
b. & {etol}         & {etoller}  & `fur stole'   & \citep{Goksel2005} \\
   & {saat}         & {saatler}  & `hour, clock' 	                  \\
%  & {harf}         & {harfler}  & `(alphabetic) letter' \\ 
\end{tabular}
\end{example}

\noindent
There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler} though their final root vowel is [$+$\textsc{Back}]. While suffix disharmony appears to very sporadic, the co-occurrence of suffix harmony and root disharmony has been the subject of much debate. As disharmonic roots still trigger suffix harmony, 
%\citet[212, 289]{Anderson1974} and
\citet{Iverson1978} distinguish suffix harmony alternations from the sequence structure constraint governing root harmony.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process. \citet[171f.]{Howard1972} correctly observes that this duplication is unnecessary.}

The disadvantage of this account is that it introduces ``duplication'' (in the sense of \citealt{Kisseberth1970b}) of sequence structure and phonological generalizations, differing only in their pattern of exceptionality. \citet[][197f.]{Zonneveld1978} shows that the theory of exceptionality proposed in \emph{SPE} can in fact account for suffix harmony in disharmonic roots.\footnote{An isomorphic analysis is available within less restrictive theories of phonological exceptionality \citep[e.g.,][]{Kisseberth1970,Pater2009}.} \citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention. A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. In other words, no form is never truly an ``exception''; rather, some forms fail to match the extended structural description of $R$ which requires that the target be [$+R$]. If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement. Thus \textsc{Backness Harmony} is able to account for both root and suffix harmony, and duplication is unnecessary.

It is important to dispense with an alternative proposed by \citet{Clements1982} and \citet{Inkelas1997} which reintroduces this duplication. Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs.  \emph{kul} `servant', \emph{kepek} `bran' vs. \emph{kapak} `lid'), whereas harmonic roots are those in which the backness of any remaining vowels is predictable. \citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots are fully specified. This is schematized below.

@@

One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, one option being a Structure Preservation condition \citep{Kiparsky1985}. However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default. A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone. However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.

While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show that Turkish speakers internalize the tendency for roots to conform to backness harmony. If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest  that this is not the case, with the caveat that uncritical acceptance of hypotheses linking loanword adaptation and language games to the grammar is no way being implied.

The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} While some speakers are said to be capable of pronouncing these non-native clusters, in fast speech the cluster is reportedly split by anaptyxis. In most cases, this vowel matches the following root vowel for backness.

\begin{example}[Variable non-native cluster adaptation (\citealp{Clements1982}:247)] 
\begin{tabular}{lllll}
a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
   & {fren}    & \alt{} & {firen}    & `brake'     \\
b. & {trablus} & \alt{} & {tırablus} & `Tripoli'   \\
   & {kral}    & \alt{} & {kıral}    & `king'      \\
c. & {brom}    & \alt{} & {burom}    & `bromide'   \\
   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
%b. & grip    & \alt{} & gırip    & `grippe'    \\ % unexpectedly back
%   & kredi   & \alt{} & kıredi   & `credit'    \\
\end{tabular}
\end{example}

\noindent
It is worth noting that this cannot be merely the consequence of application of the native \textsc{Backness Harmony} rule, since the leftmost vowel of a prosodic word is not subject to harmonization. Rather, the adaptation of non-native onset clusters proceeds in such a fashion so that the lexical items in question are rarely [$-$\textsc{Backness Harmony}].

Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} 
The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults. In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel. In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the the inserted [$+$\textsc{Back}] vowel. The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}

@@

\noindent 
\citeauthor{Harrison2001} present an analysis of this game couched in the vowel underspecification proposed by \citeauthor{Clements1982}, but the data is equally consistent with a full specification analysis. Reharmonization is the result of \textsc{Backness Harmony} in the reduplicant. On the other hand, the lack of reharmonization in the disharmonic roots indicates that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.

A number of studies have investigated the role of harmony in word-spotting tasks, though to mimic auditory word recognition and segmentation in natural settings. Many of these studies have been carried out in Finnish, which has a vowel harmony system similar to that of Turkish. \citet{Suomi1997} and \citet{Vroomen1998} task Finnish speakers with identifying harmonic disyllables in an auditory stream. When the syllable preceding the disyllabic target has a different backness specification than the target, recognition of the target is facilitated. Presumably, disharmony facilitates the recognition of word boundaries. \citet{Kabak2010} find that Turkish \textsc{Backness Harmony} has a a similar effect: Turkish speakers are quicker and more accurate at the task of spotting the nonce target word \emph{pavo} when preceded by a disharmonic juncture (e.g., \emph{gölü-PAVO}) than when preceded by a harmonic juncture (e.g., \emph{golu-PAVO}). \citet{Kabak2010} find that this effect does not obtain for speakers of French, a language which lacks vowel harmony. This implies that Turkish speakers have internalized the tendency of harmonic sequences to be root-internal and of disharmonic transitions to cross word boundaries.

>>>>

There is a long tradition of distinguishing between structural phonotactic gaps and those which are simply accidental \citep[e.g.,][]{Fischer-Jorgensen1952,Saporta1955,Saporta1958,Vogt1954}. In purely statistical terms, an accidental gap is one which is likely to arise without any antecedent cause, i.e., by chance. The contribution of early generative discussions of this contrast \citep[e.g.,][]{Halle1962,Chomsky1965,SPE} is to ground the definitions of \emph{structural} and \emph{accidental} gaps not in statistical expectations, but rather in mentalism. For \citeauthor{SPE}, an accidental gap is one which is not learned, one which lacks a mental antecedent.

Much of the latter-day literature on phonotactic knowledge, while adopting the representational machinery of generative phonology, takes less interest in the question of what generalizations speakers internalize. In a seminal paper, \citet{McCarthy1988} uses a simple statistical technique to validate a generalization about Arabic root consonant co-occurrence facts originally due to \citet{Greenberg1950}. The assumption appears to be that static co-occurrence restrictions which have a statistically significant trace in the lexicon will also be reflected in psycholinguistic tasks such as wordlikeness judgements. While the Arabic restriction discussed by \citeauthor{McCarthy1988} was ultimately investigated experimentally \citep{Frisch2004}, this is the case for only a small minority of the studies inspired by \citeauthor{McCarthy1988}'s study; in most cases, phonotactic generalizations are inferred directly from lexical statistics \citep[e.g.,][]{Anttila2008a,Berkley1994b,Berkley1994a,Berkley2000,Brown2010,Buckley1997,Coetzee2008a,Elmedlaoui1995,Graff2011,MacEachern1999,Kinney2005,Kawahara2006,Martin2007,Martin2011,Mester1988,Miller-Ockhuizen2003,Padgett1992,Padgett1995,Pozdniakov2007,Yip1989}. In a study of consonant co-occurrence restrictions in Gitksan, \citet{Brown2010} claims that \emph{any} statistically significant pattern in the lexicon is one that is internalized by speakers:

@@

There are several arguments against this inference. First, negative results in formal learnability theory (see \citealt{Yang2012} for a recent review) show that there are whole classes of rules and constraints which cannot be learned under standard assumptions about acquisition (e.g., that the data is finite, can appear in any order, and that speakers do not have to request grammatical judgements from adults to achieve adult-like competence).

Secondly, there are a number of results which statistically reliable patterns are not learned by speakers \citep[e.g.,][]{Becker2011,Hayes2006,Hayes2009,HayesInPress}. This argues against a direct link between statistical criteria and the synchronic reality of phonotactic generalizations. The danger is that grammatical models which closely mimic lexical statistics but which lack psycholinguistic support (e.g., Gitksan: \citealt{Brown2010}; Muna: \citealt{Anttila2008a}, \citealt{Coetzee2008a}; Shona: \citealt[][385]{Hayes2008a}) may overstate the phonotactic generalizations internalized by speakers.

Another type of non-learning occurs when generalizations are statistically reliable but demonstrably incorrect. \citet{Legate2012}, for instance, note that the overwhelming majority of English words have initial primary stress. While this suggests a quantity-insensitive stress system, English stress is in fact quantity-sensitive (see \citealt{Halle1998c} and references therein).

Finally, there are many cases of statistical generalizations with a known diachronic cause but lacking any synchronic reality. One of the earliest arguments of this type was put forth by Saussure (\citeyear[202f.]{CLG}) in a discussion of rhotacism, a diachronic process which changed Old Latin intervocalic \emph{s} to \emph{r}. As a result, Classical Latin intervocalic \emph{s} is rare. However, Saussure denies that this generalization has any status in Classical Latin.\footnote{See \citet{Gorman2012e} and citations therein for a full explication of Saussure's position on rhotacism, and a restatement of the phonological argument in generative terms.}

Similarly, [ʃ] in Modern English derives from Old English [sk] (e.g., \emph{fisc} `fish'), and since long vowels before complex codas are not found in Old English, [ʃ] is rarely preceded by long vowels in word-final syllables to this day according to \citet{Iverson2005}. The type frequencies of short and long vowels before final [ʃ] and a similar segment, [s] are shown in Table \ref{ssh}. As can be seen, long vowels before final [s] are twice as common before final [ʃ]; according to the appropriate statistical test (the Fisher exact test, described below), this is significant. Yet \citet{Iverson2005} label this constraint ``phonologically accidental'' since a millennium of speakers have produced coinages (e.g., \emph{posh}, \emph{swoosh}) and  borrowings (e.g., \emph{douche}, \emph{capiche}) that disregard this generalization.\footnote{In the loanword adaptation literature, there are many reports that statistically reliable generalizations over the native lexicon are ignored in nativization processes \citep[e.g.,][]{Ito1995a,Ito1995b,Ussishkin2003}. On the other hand, \citet{Frisch2001} suggest that novel borrowings in Egyptian Arabic rarely violate co-occurrence restrictions in that language; however, the generalization in question was exceptionless from Proto-Semitic \citep{Ehret1989} through Classical Arabic \citep{Greenberg1950}, so any violations of this generalization are unexpected.}

\begin{table}
\centering
\begin{tabular}{l r r r r}
\toprule
          & \{ɪ, ɛ, æ, ʌ, ʊ\}\gap{}\# & \{i, e, ɑ, ɔ, u\}\gap{}\# & \% long & $p$-value \\
\midrule
\gap{}ʃ\# & 78                & 9                 & 8\%      & \multirow{2}{*}{0.026} \\
\gap{}s\# & 410               & 107               & 16\%     & \\
\bottomrule
\end{tabular}
\caption{Type frequencies of short and long vowels before word-final [ʃ] and [s] in a subset of the CMU pronunciation dictionary containing all words with token frequency greater than 1 per million words in the SUBTLEX-US frequency norms; the 9 words containing ending in a long vowel followed by ʃ are \emph{douche}, \emph{leash}, \emph{gosh}, \emph{josh}, \emph{posh}, \emph{squash}, \emph{unleash}, \emph{wash}, and \emph{woosh}.}
\label{ssh}
\end{table}

This chapter is a comprehensive investigation of three phonotactic generalizations in Turkish, comparing lexical statistics and the results of a wordlikeness task performed by \citet{Zimmer1969}. Both the lexical statistics and \citeauthor{Zimmer1969}'s experimental results merit reconsideration, because prior discussions do not relate the lexical statistics to competing formalizations of the generalizations involved, and there is widespread disagreement concerning the experimental findings.

One of the three generalizations, \textsc{Labial Attraction}, is shown to   state a statistically reliable generalization about the Turkish lexicon, but has no effect on speakers' wordlikeness judgements. This further undermines the widespread use of statistical criteria to determine which of the vast number of statistically reliable phonotactic generalizations are internalized by speakers and which are not. The simplest hypothesis is that \textsc{Labial Attraction} is inactive in wordlikeness tasks because it is static, corresponding to no phonological process in the language.

@@

\citet{Lees1966b,Lees1966a} proposes three constraints on Turkish vowel sequences; these constraints are the focus of many subsequent studies. In this section, these constraints are formalized, and where possible, related to phonological alternations and to behavioral evidence bearing on speakers' knowlege of the restrictions. The following feature specification for the eight vowels of Turkish is assumed throughout.

@@

Two less familiar notations are used in this chapter. First, directional application conditions are assumed, so as to derive the left-to-right, iterative properties of the harmony rules. Directional application provides a panacea for the considerable body of evidence against simultaneous rule application adduced in the 1970s (e.g., \citealt[209f.]{Anderson1974}, \citealt{Howard1972}, \citeauthor{Kenstowicz1973} \citeyear[14f.]{Kenstowicz1973}, \citeyear[189f.]{Kenstowicz1977}, \citeyear[318f.]{Kenstowicz1979}, \citealt{Piggott1975}). Further, \citet{Johnson1972} and \citet{Kaplan1994} prove directional application and simultaneous application are equivalent in terms of formal learnability.

Secondly, rather than the use of an unbounded number of Greek-letter variables ($\alpha$, $\beta$, etc.) over feature values $\{+, -\}$, only a single variable, denoted by `$=$', is used \citep{McCawley1973}. A structural description [$=$F]\ldots{}[$=$F] matches a string S$_i$\ldots{}S$_j$ if and only if S$_i$ and S$_j$ are both [$+$F] or both [$-$F]. This is more restrictive than Greek-letter notation, in that it prevents the value of one feature being applied to another. \citet{Odden2012} argues that the two counterexamples against this restriction adduced in \emph{SPE} (352-353) are not probative.

@@

If permitted to apply in non-derived environments, this rule accounts for the tendency of polysyllabic roots to contain only [$+$\textsc{Back}] or [$-$\textsc{Back}] vowels, a tendency which will be quantified below. \textsc{Backness Harmony} also triggers alternations in inflectional suffix vowels. For instance, the nominative plural (nom.pl.) suffix is \emph{-ler} when the final root vowel is [$-$\textsc{Back}], and \emph{-lar} when it is [$+$\textsc{Back}].

\begin{example}[The Turkish nominative]
\label{turknom}
\begin{tabular}{lllll}
   & \emph{nom.sg.} & \emph{nom.pl.} \\
a. & {ip}           & {ipler}    & `rope'         & \citep[][216]{Clements1982} \\
   & {köy}          & {köyler}   & `village'      \\
   & {yüz}          & {yüzler}   & `face'         \\
   & {kız}          & {kızlar}   & `girl'         \\
   & {pul}          & {pullar}   & `stamp'        \\
b. & {neden}        & {nedenler} & `reason'       & \citep{TELL} \\
   & {kiler}        & {kilerler} & `pantry'       \\
   & {pelür}        & {pelürler} & `onionskin'    \\
   & {boğaz}        & {boğazlar} & `throat'       \\
   & {sapık}        & {sapıklar} & `pervert'      \\
\end{tabular}
\end{example}

A few complications arise, however. First, as shown in (\ref{turkexcept}a), not all polysyllabic roots conform to \textsc{Backness Harmony}. In this case, suffix vowels generally exhibit harmony with the final root vowel. There is also a very small class of nouns, shown in (\ref{turkexcept}b), which take \emph{-ler}, although their final root vowel is [$+$\textsc{Back}]. Interestingly, they may themselves be harmonic.

\begin{example}[Exceptional Turkish nominatives] 
\label{turkexcept}
\begin{tabular}{lllll}
   & \emph{nom.sg.} & \emph{nom.pl.}&                    \\
a. & {mezar}        & {mezarlar}    & `grave'       & \citep{TELL}       \\
   & {model}        & {modeller}    & `model'                            \\
   & {silah}        & {silahlar}    & `weapon'                           \\
   & {memur}        & {memurlar}    & `official'                         \\
   & {sabun}        & {sabunlar}    & `soap'                             \\
b. & {etol}         & {etoller}     & `fur stole'   & \citep{Goksel2005} \\
   & {saat}         & {saatler}     & `hour, clock' 	              \\
   & {kahabat}      & {kahabatler}  & `fault'       \\
\end{tabular}
\end{example}

\citet[212]{Anderson1974} and \citet{Iverson1978} argue that suffix harmony in disharmonic roots found in (\ref{turkexcept}a) requires the rule governing suffix harmony alternations to be distinguished from a sequence structure constraint governing root harmony. Since the rule and sequence structure constraint are otherwise identical, this constitutes an undesirable ``duplication'' in the sense of \citealt{Kisseberth1970b}. However, the theory of exceptionality proposed in \emph{SPE} can account for these facts without duplication \citep[197f.]{Zonneveld1978}.\footnote{\citet[29f.]{Kiparsky1968} discusses a parallel case in Finnish, and proposes a similar separation between the exception-filled sequence structure constraint and an exceptionless suffix harmony process. \citet[171f.]{Howard1972} correctly observes that this duplication is unnecessary.}

\citeauthor{SPE} assume that the specification of the target (i.e., the segment or segments to be changed) of a rule \emph{R} must be marked [$+$\emph{R}] by convention. A root or affix which fails to undergo \emph{R} despite otherwise matching the structural description is simply said to be marked [$-$\emph{R}]. In other words, no form is never truly an ``exception''; rather, some underlying representations have non-default features which do not match the extended structural description of $R$ which requires that the target be [$+R$]. If disharmonic roots are marked [$-$\textsc{Backness Harmony}], then the final vowel of disharmonic roots will still trigger \textsc{Backness Harmony}, since the [$-$\textsc{Backness Harmony}] root is no longer the target but rather the trigger, which is not subject to the [$+$\textsc{Backness Harmony}] requirement.

Under this account, root and suffix harmony are derived from \textsc{Backness Harmony}, but they do not have the same ontology: suffix harmony is direct result of phonological rule application, whereas any ``effects'' of the  generalization concerning harmonic roots arises from a dispreference for lexical exceptionality.

\citeauthor{Anderson1974} also notes that roots which fail to undergo suffix harmony, like (\ref{turkexcept}b), may themselves be harmonic. He takes to be evidence for the necessity of duplication:

\begin{quote}
\ldots{}there are words which are exceptions to harmony across boundaries (e.g., \emph{kabahat} `fault', \emph{kabahattı} `his fault') but which are perfectly regular internally. Since the morpheme structure condition and the phonological rule in this case have distinct classes of exceptions, it is clear that they cannot be identified. \citep[289]{Anderson1974}
\end{quote}

\noindent
Under the assumptions so far, there is no reason to think that \emph{kabahat} is [$+$\textsc{Backness Harmony}], however: even roots with all back vowels need not undergo this rule. Whatever the ultimate analysis of those few words which fail to show suffix harmony, nothing depends on whether or not they are harmonic.

%\begin{example}[Turkish progressives]
%\begin{tabular}{l l l l l}
%   & \emph{1sg. present}  & \emph{1sg. progressive} & \\
%a. & gelirim              & geliyorum               & `come'  \\
%b. & görürüm              & görüyorum               & `see'   \\
%c. & atarım               & atıyorum                & `throw' \\
%d. & bulurum              & buluyorum               & `find'  \\
%\end{tabular}
%\end{example}

It is necessary to dispense with an alternative analysis proposed by \citet{Clements1982} and \citet{Inkelas1997}. This analysis has desirable properties, but a subtlety of it reintroduces the duplication of rule and sequence structure constraint. Root vowels exhibit a robust contrast for backness (e.g., \emph{kül} `ash' vs. \emph{kul} `servant', \emph{kepek} `bran' vs. \emph{kapak} `lid'), whereas backness of vowels in non-initial syllables is predictable in harmonic roots; there are no prefixes in Turkish. \citeauthor{Clements1982} propose that these vowels, as well as harmonizing suffix vowels, are underspecified for backness, whereas the non-initial vowels of disharmonic roots and of certain exceptional suffixes are fully specified. This is schematized below.

@@

One crucial detail is missing from this analysis: \textsc{Backness Harmony} needs to be prevented from overwriting the [$+$\textsc{Back}] specification of disharmonic roots, either by non-derived environment blocking \citep{Mascaro1976} or structure preservation condition \citep{Kiparsky1985}. However, any condition which prevents \textsc{Backness Harmony} from overwriting underlying backness specifications will reintroduce the duplication of sequence structure and phonological generalizations; under this analysis, disharmonic roots are no longer exceptional, despite considerable evidence (reviewed below) that they are marked in Turkish.\footnote{On the other hand, it is possible to interpret the presence of a single backness specification per root as a sort of default. A precedent for this is the surface-oriented interpretation of the tonal Obligatory Contour Principle proposed by \citet[134]{Goldsmith1976} and \citet{Odden1986}, under which adjacent identical tones are automatically attributed to a single underlying tone. However, this is merely a notational variant of the rule exceptionality account in which [$+$\textsc{Backness Harmony}] is the default.} On these grounds, the underspecification analysis is rejected here in favor of the rule exceptionality account.

While harmony in non-derived environments can be inferred from the aforementioned suffix alternations, no evidence has yet been presented to show that Turkish speakers internalize the tendency for roots to conform to backness harmony. If Turkish speakers do not attend to this generalization, there is no need for the grammar to account for it. Several other ``external'' facts suggest that this is not the case. The discussion here is not intended to imply uncritical acceptance of evidence from loanword adaptation, language games or psycholinguistic tasks as evidence for phonological grammar, but rather to illustrate additional evidence that is pertinent if the linking hypothesis is correct.

The production of non-native word-initial onset clusters, discussed by \citet{Clements1982} and \citet{Kaun1999}, suggests that loanword adaptation respects \textsc{Backness Harmony}.\footnote{Thanks to Kie Zuraw for bringing this data to my attention.} Some speakers are said to be capable of pronouncing these non-native clusters, but in fast speech the cluster is split by anaptyxis. In most cases, this vowel matches the following root vowel for backness.

\begin{example}[Variable non-native cluster adaptation (\citealp{Clements1982}:247)] 
\begin{tabular}{lllll}
a. & {spiker}  & \alt{} & {sipiker}  & `announcer' \\
   & {fren}    & \alt{} & {firen}    & `brake'     \\
b. & {trablus} & \alt{} & {tırablus} & `Tripoli'   \\
   & {kral}    & \alt{} & {kıral}    & `king'      \\
c. & {brom}    & \alt{} & {burom}    & `bromide'   \\
   & {prusya}  & \alt{} & {purusya}  & `Prussia'   \\
\end{tabular}
\label{spiker}
\end{example}

\noindent
It is unclear whether the cluster-splitting vowel is deleted in the non-native variant or epenthesized in the fast speech variant. Under either analysis, there is no ready explanation for the tendency of the cluster-splitting vowel to have the backness features of following consonants; if anything, one might expect it to determine the backness features of following consonants. All that can be said with certainty is that the adaptation of non-native onset clusters appears to proceed in such a fashion so that the lexical items in question are [$-$\textsc{Backness Harmony}].

Similar evidence comes from a language game discussed by \citet{Harrison2001}.\footnote{Thanks to Bert Vaux for bringing this study to my attention.} 
The game is native to the related language Tuvan, where it is used to convey a sense of ``vagueness or jocularity''; it is not indigenous to Turkish, but can be taught quickly to children or adults. In this game, the base is reduplicated and the first vowel of the reduplicant replaced with a [$+$\textsc{Back}] vowel. In (\ref{redupgame}a), the second [$-$\textsc{Back}] vowel of the base is, in the reduplicant, ``reharmonized'' with the inserted [$+$\textsc{Back}] vowel. The disharmonic roots of (\ref{redupgame}b) do not reharmonize.\footnote{A similar contrast between harmonic and disharmonic roots is found in Tuvan \citep{Harrison2001} and in an unrelated Finnish language game \citep{Campbell1986}.}

@@

\noindent 
In the full specification analysis adopted here, reharmonization is the result of \textsc{Backness Harmony} applying within the reduplicant. On the other hand, the lack of reharmonization in the reduplicants of disharmonic roots suggests that the [$-$\textsc{Backness Harmony}] exception feature is copied under reduplication.\footnote{The strongest argument that reduplication preserves diacritic features of the root would be a case where lexical exceptionality surfaces only in reduplicants. Kinande reduplication, documented by \citet{Mutaka1990} and further discussed by \citet{Downing2000}, provides such a case. Verb reduplication requires a bisyllabic reduplicant. This clearly has synchronic force in the gramar, since reduplicated monosyllabic roots actually have three copies of the root, and since reduplicated forms of many trisyllable verbs are ineffable. However, a few trisyllabic verbs exceptionally show full reduplication. Exceptionality appears to be a lexical property (since it is not predictable), but it surfaces only in the reduplicant.} \citet{Silverman2000} notes that static phonotactic generalizations are not preserved by operations like reduplication and truncation: under this definition, root harmony is not static.
